---
title: "Deep_Neural_Networks_With_R"
author: "Gino Tesei"
date: "February 25, 2016"
output:
  html_document:
    highlight: tango
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document:
    highlight: tango
    number_sections: yes
    toc: yes
---

# How to Immediately Approximate Any Function

__Hornik et al. theorem__ 
Let F be a continuous function on a bounded subset of n-dimensional space. Then there exists a two-layer neural network F with a finite number of hidden units that approximate F arbitrarily well. Namely, for all x in the domain of F, $\begin{equation}|F(x)−\hat{F}(x)|<\epsilon.\end{equation}.$

```{r,warning=F,message=FALSE,echo=TRUE}
Rsquared <- function(obs,preds) {
  mobs = mean(obs)
  1-sum((obs-preds)^2)/sum((obs-mobs)^2)
}

library ("neuralnet")
require (Metrics)
set.seed (2016) 
attribute <- as.data.frame( sample(seq (-2 ,2, length =50), 50, replace = FALSE) , ncol =1) 
response <-attribute ^2
data <- cbind( attribute, response) 
colnames (data) <- c( "attribute","response")
head (data ,10)
with(data,plot(attribute,response,color="gray",xlab="attribute",ylab="response"))

##
fit <-neuralnet (response ~ attribute, data=data, hidden =c(3,3), threshold = 0.01)
library(devtools)
source_url('https://gist.githubusercontent.com/fawda123/7471137/raw/466c1474d0a505ff044412703516c34f1a4684a5/nnet_plot_update.r')
plot.nnet(fit)

testdata <- as.matrix( sample(seq( -2 ,2, length =10) , 10, replace = FALSE) , ncol =1)
pred <- compute (fit, testdata)
result <- cbind( testdata, pred$net.result , testdata^2) 
colnames(result) <- c( " Attribute " ," Prediction ", " Actual ") 
round(result,4)
rmse(actual = testdata^2 , predicted = pred$net.result)
Rsquared(obs=testdata^2,preds=pred$net.result)

plot(x=testdata,y=testdata^2,col="red",xlab="testdata",ylab="pred vs. actual")
points(x=testdata, y=pred$net.result,col="blue") 
```

# The Boston dataset 

```{r,warning=F,message=FALSE,echo=TRUE}
data( "Boston" ,package = "MASS") 
data <-Boston
apply(X = data,MARGIN = 2,FUN = function(x) sum(is.na(x)))
caret::featurePlot(x = data[,-grep(pattern = "medv",x = colnames(data))], y = data$medv,
            between = list(x = 1, y = 1),
            type = c("g", "p", "smooth")) 

##
f<-medv~ crim + indus + nox + rm + age + dis + tax + ptratio + lstat
set.seed(2016) 
n =nrow(data) 
train <- sample (1:n, 400, FALSE)

## neuralnet
# fit <- neuralnet (f, data = data[ train,], hidden =c(10 ,12 ,20), 
#                   algorithm = "rprop+" , 
#                   err.fct = "sse" , 
#                   act.fct = "logistic" , 
#                   threshold = 0.1, 
#                   linear.output = TRUE)
# 
# pred <- compute(fit, data[ -train , 1:9])
# 
# ##
# round(cor( pred$net.result, data[ - train ,10]) ^2 ,6)
# mse(data [-train ,10], pred$net.result)
# rmse(data [-train ,10], pred$net.result) 

## 
# require(deepnet)
# set.seed (2016) 
# X = data[ train ,1:9] 
# Y = data[ train ,10] 
# fitB <-nn.train (x = as.matrix(X), y =Y, 
#                   initW = NULL, 
#                  initB = NULL, 
#                  hidden = c(10 ,12 ,20), 
#                  learningrate = 0.58, 
#                  momentum = 0.74, 
#                  learningrate_scale = 1,
#                  activationfun = "sigm" , 
#                  output = "linear" , 
#                  numepochs = 970, 
#                  batchsize = 60, 
#                  hidden_dropout = 0, 
#                  visible_dropout = 0)
# 
# Xtest <- data[ -train ,1:9] 
# predB <- nn.predict (fitB, Xtest)
# 
# round(cor( predB ,data[ -train ,10]) ^2 ,6) 
# mse (data [-train ,10], predB) 
# rmse (data [-train ,10], predB) 
```

# Binary Classification problems 

```{r,warning=F,message=FALSE,echo=TRUE}
data( "PimaIndiansDiabetes2" ,package = "mlbench")
ncol(PimaIndiansDiabetes2)
nrow(PimaIndiansDiabetes2)

# NAs
apply(X = PimaIndiansDiabetes2,MARGIN = 2,FUN = function(x) sum(is.na(x)))
temp <-(PimaIndiansDiabetes2) 
temp $insulin <- NULL 
temp $triceps <- NULL 
temp <-na.omit( temp)
nrow( temp)

# 
y<-( temp $diabetes) 
temp $diabetes <-NULL 
temp <- scale( temp) 
temp <- cbind(as.factor(y), temp)
class(temp)
summary(temp)

# 
set.seed (2016) 
n =nrow( temp) 
n_train <- 600 
n_test <-n - n_train 
train <- sample (1: n, n_train, FALSE)

require (RSNNS)
set.seed (2016) 
X<-temp [train ,1:6] 
Y<-temp [train ,7]

fitMLP <- mlp (x =X, y =Y, size = c(12 ,8), maxit = 1000, 
               initFunc = "Randomize_Weights" , initFuncParams = c( -0.3, 0.3), 
               learnFunc = "Std_Backpropagation" , 
               learnFuncParams = c(0.2, 0), 
               updateFunc = "Topological_Order " , 
               updateFuncParams = c(0), 
               hiddenActFunc = "Act_Logistic" , 
               shufflePatterns = TRUE, linOut = TRUE)


predMLP <- sign( predict (fitMLP, temp [- train ,1:6]))

table( predMLP ,sign( temp [-train ,7]), 
       dnn =c( "Predicted", "Observed"))

error_rate = (1 - sum( predMLP == sign( temp [-train ,7]))/ 124) 
round( error_rate ,3)

# AMORE
detach( "package:RSNNS", unload = TRUE) 
library (AMORE)

net <- newff (n.neurons =c(6 ,12 ,8 ,1), 
              learning.rate.global =0.01, 
              momentum.global =0.5, 
              error.criterium = "LMLS" , 
              Stao = NA, 
              hidden.layer = "sigmoid" , 
              output.layer = "purelin" , 
              method = "ADAPTgdwm")

X<-temp [train,-7] 
Y<-temp [train ,7]

fit <- train (net, P =X , T =Y , 
              error.criterium = "LMLS" , 
              report= TRUE, show.step =100, n.shows =5)

pred <- sign( sim (fit$net, temp [-train,]))
table( pred ,sign( temp [- train ,7]), dnn =c( "Predicted" , "Observed"))


error_rate = (1 - sum( pred == sign( temp [-train ,7]))/ 124) 
round( error_rate ,3) 
```

# Multiple Response Classification problems  

```{r,warning=F,message=FALSE,echo=TRUE}
data( "bodyfat" ,package = "TH.data")

set.seed (2016) 
train <- sample (1:71 ,50, FALSE)

scale_bodyfat <-as.data.frame(scale(log(bodyfat))) 
f<- waistcirc + hipcirc ~ DEXfat + age + elbowbreadth + kneebreadth + anthro3a + anthro3b + anthro3c + anthro4

# it <- neuralnet (f, data = scale_bodyfat [train,], 
#                  hidden =c(8 ,4), threshold =0.1,
#                  err.fct = "sse" , 
#                  algorithm = "rprop+" , 
#                  act.fct = "logistic" , 
#                  linear.output = FALSE )
# 
# without_fat <- scale_bodyfat 
# without_fat$waistcirc <-NULL  
# without_fat$hipcirc <-NULL
# 
# pred <- compute (fit, without_fat [-train,] ) 
# pred $net.result

### installed packages 
# pack <- as.data.frame (installed.packages () [,c(1 ,3:4)]) 
# rownames (pack) <- NULL 
# pack <- pack [is.na( pack$Priority), 1:2, drop= FALSE] 
# print( pack, row.names= FALSE)
```

# The Elman Neural Networks 
```{r,warning=F,message=FALSE,echo=TRUE}
require (RSNNS) 
require (quantmod)

data( "UKLungDeaths" ,package = "datasets")

par( mfrow =c(3 ,1)) 
plot( ldeaths, xlab = "Year" , ylab = "Both sexes" , main = "Total") 
plot( mdeaths, xlab = "Year" , ylab = "Males" , main = "Males") 
plot( fdeaths, xlab = "Year" , ylab = "Females" , main = "Females")

sum(is.na( ldeaths))

class( ldeaths)

par( mfrow = c(3, 1))
plot( ldeaths)

x<- density (ldeaths)

plot(x, main = "UK total deaths from lung diseases") 
polygon (x, col= "green", border = "black") 
boxplot (ldeaths ,col= "cyan", ylab = "Number of deaths per month")

#
y<-as.ts( ldeaths)
y<- log( y)
y<- as.ts(scale( y)) ### ??????????????

y<-as.zoo (y)
x1 <-Lag (y, k = 1) 
x2 <-Lag (y, k = 2) 
x3 <-Lag (y, k = 3) 
x4 <-Lag (y, k = 4) 
x5 <-Lag (y, k = 5) 
x6 <-Lag (y, k = 6) 
x7 <-Lag (y, k = 7)
x8 <-Lag (y, k = 8) 
x9 <-Lag (y, k = 9) 
x10 <-Lag (y, k = 10) 
x11 <-Lag (y, k = 11) 
x12 <-Lag (y, k = 12)

deaths <- cbind( x1, x2, x3, x4, x5, x6, x7, x8, x9, x10 ,x11, x12) 
deaths <- cbind(y, deaths)
deaths <- deaths [-(1:12),]
n =nrow( deaths)
n

set.seed(465)
n_train <- 45 
train <- sample (1:n, n_train, FALSE)

inputs <- deaths [,2:13] 
outputs <- deaths [,1]

fit <- elman (inputs [train], 
                outputs [train], 
                size =c(1 ,1), 
                learnFuncParams =c(0.1), 
                maxit =1000)

plotIterativeError (fit)

summary(fit)

#
pred <- predict (fit, inputs [-train])

cor( outputs [-train], pred)^2

rmse(actual = outputs [-train] , predicted = pred)
Rsquared(obs = outputs [-train]  , preds = pred)
```

# The Jordan Neural Networks 
```{r,warning=F,message=FALSE,echo=TRUE}
require (RSNNS) 
data( "nottem" ,package = "datasets") 
require (quantmod)

class( nottem)
plot( nottem)

#
y<-as.ts( nottem) 
y<- log( y) 
y<- as.ts(scale( y))

y<-as.zoo (y) 
x1 <-Lag (y, k = 1) 
x2 <-Lag (y, k = 2)
x3 <-Lag (y, k = 3) 
x4 <-Lag (y, k = 4) 
x5 <-Lag (y, k = 5) 
x6 <-Lag (y, k = 6) 
x7 <-Lag (y, k = 7) 
x8 <-Lag (y, k = 8) 
x9 <-Lag (y, k = 9) 
x10 <-Lag (y, k = 10) 
x11 <-Lag (y, k = 11) 
x12 <-Lag (y, k = 12)


# 
temp <- cbind( x1, x2, x3, x4, x5, x6, x7, x8, x9, x10 ,x11, x12) 
temp <- cbind(y, temp) 
temp <- temp [-(1:12),]

plot( temp)

# 
n =nrow(temp) 
n 
set.seed (465)
n_train <- 190 
train <- sample (1:n, n_train, FALSE)

# 
inputs <- temp [,2:13] 
outputs <- temp [,1] 
fit <- jordan (inputs [train], 
               outputs [train], 
               size =2, 
               learnFuncParams =c (0.01), 
               maxit =1000)

plotIterativeError(fit)

pred <- predict (fit, inputs [-train]) 
cor( outputs [-train], pred)^2 

rmse(actual = outputs [-train] , predicted = pred)
Rsquared(obs = outputs [-train]  , preds = pred)
```