---
title: "12_Using_Colors"
author: "Gino Tesei"
date: "December 14, 2015"
output: html_document
---

## 1. Setting the Colors of Objects
In the call to the geom, set the values of colour or fill
```{r}
library(ggplot2)
library(gcookbook) # For the data set 

library(plyr)
library(scales)

ggplot( mtcars, aes( x = wt, y = mpg)) + 
  geom_point( colour ="red") 

library( MASS) # For the data set 
ggplot( birthwt, aes( x = bwt)) + 
  geom_histogram( fill ="red", colour ="black")
```

## 2. Mapping Variables to Colors
You want to use a variable (column from a data frame) to control the color of geoms.
```{r}
ggplot( cabbage_exp, aes( x = Date, y = Weight, fill = Cultivar)) + 
  geom_bar( colour ="black", position ="dodge" , stat = "identity") 

ggplot( mtcars, aes( x = wt, y = mpg, colour = cyl)) + 
  geom_point()

# Convert to factor in call to ggplot() 
ggplot( mtcars, aes( x = wt, y = mpg, colour = factor( cyl))) + 
  geom_point() 
```

## 3. Using a Different Palette for a Discrete Variable
```{r}
# Base plot 
p <- ggplot( uspopage, aes( x = Year, y = Thousands, fill = AgeGroup)) + 
  geom_area() 
p 

# ColorBrewer palette 
p + scale_fill_brewer()

# ColorBrewer palette Oranges
p + scale_fill_brewer( palette ="Oranges")

# gray 
p + scale_fill_grey()

# Reverse the direction and use a different range of greys 
p + scale_fill_grey( start = 0.7, end = 0)
```

## 4. Using a Manually Defined Palette for a Discrete Variable
```{r}
# Base plot 
h <- ggplot( heightweight, aes( x = ageYear, y = heightIn, colour = sex)) + 
  geom_point() 

h

# Using color names 
h + scale_colour_manual( values = c("red", "blue")) 

# Using RGB values 
h + scale_colour_manual( values = c("#CC6666", "#7777DD"))
```

## 5. Using a Colorblind-Friendly Palette
```{r}
# Base plot 
p <- ggplot( uspopage, aes( x = Year, y = Thousands, fill = AgeGroup)) + 
  geom_area() 

# The palette with grey: 
cb_palette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") 

# Add it to the plot 
p + scale_fill_manual( values = cb_palette)
```

## 6. Using a Colorblind-Friendly Palette
```{r}
# Base plot
p <- ggplot( heightweight, aes( x = ageYear, y = heightIn, colour = weightLb)) + 
  geom_point( size = 3) 
p 

# With a gradient between two colors 
p + scale_colour_gradient( low ="black", high ="white") 

# A gradient with a white midpoint library( scales) 
p + scale_colour_gradient2( low = muted("red"), mid ="white", high = muted("blue"), midpoint = 110) 

# A gradient of n colors 
p + scale_colour_gradientn( colours = c("darkred", "orange", "yellow", "white"))
```

## 7. Coloring a Shaded Region Based on Value
```{r}
cb <- subset( climate, Source =="Berkeley") 
cb$valence[ cb$Anomaly10y >= 0] <- "pos" 
cb$valence[ cb$Anomaly10y < 0] <- "neg" 
str(cb)

ggplot( cb, aes( x = Year, y = Anomaly10y)) + 
  geom_area( aes( fill = valence)) + 
  geom_line() + 
  geom_hline( yintercept = 0)


# approx() returns a list with x and y vectors 
interp <- approx( cb$Year, cb$Anomaly10y, n = 1000) 

# Put in a data frame and recalculate valence 
cbi <- data.frame( Year = interp$x, Anomaly10y = interp$y) 
cbi$valence[ cbi$Anomaly10y >= 0] <- "pos" 
cbi$valence[ cbi $ Anomaly10y < 0] <- "neg"


## plot with intrapolations cbi 
ggplot( cbi, aes( x = Year, y = Anomaly10y)) + 
  geom_area( aes( fill = valence), alpha = .4) +
  geom_line() + geom_hline( yintercept = 0) + 
  scale_fill_manual( values = c("#CCEEFF", "#FFDDDD"), guide = FALSE) + 
  scale_x_continuous( expand = c( 0, 0))
```

