---
title: "3_Bar_Graphs"
author: "Gino Tesei"
date: "December 12, 2014"
output: html_document
---

## 1. Making a Basic Bar Graph
```{r}
library(ggplot2)
library( gcookbook) # For the data set 

library(plyr) ## 


ggplot( BOD, aes( x = Time, y = demand)) + geom_bar( stat ="identity")
ggplot( BOD, aes( x = factor( Time), y = demand)) + geom_bar( stat ="identity")
ggplot( pg_mean, aes( x = group, y = weight)) + geom_bar( stat ="identity", fill ="lightblue", colour =" black") 
```


## 2. Grouping Bars Together
```{r}
ggplot( cabbage_exp, aes( x = Date, y = Weight, fill = Cultivar)) + 
  geom_bar( position ="dodge", stat ="identity")

ggplot( cabbage_exp, aes( x = Date, y = Weight, fill = Cultivar)) + 
  geom_bar( position ="dodge", colour ="black", stat ="identity") + 
  scale_fill_brewer( palette ="Pastel1")
```

## 3. Making a Bar Graph of Counts
```{r}
ggplot( diamonds, aes( x = cut)) + geom_bar()
```

## 4. Using Colors in a Bar Graph
```{r}
upc <- subset( uspopchange, rank( Change) > 40)

ggplot( upc, aes( x = Abb, y = Change, fill = Region)) + geom_bar( stat ="identity")

ggplot( upc, aes( x = reorder( Abb, Change), y = Change, fill = Region)) + 
  geom_bar( stat ="identity", colour ="black") + 
  scale_fill_manual( values = c("#669933", "#FFCC66")) + xlab("State")

```

## 5. Coloring Negative and Positive Bars Differently
```{r}
csub <- subset( climate, Source =="Berkeley" & Year >= 1900) 
csub$pos <- csub$Anomaly10y >= 0
head(csub)

ggplot( csub, aes( x = Year, y = Anomaly10y, fill = pos)) + 
  geom_bar( stat ="identity", position ="identity") 

ggplot( csub, aes( x = Year, y = Anomaly10y, fill = pos)) + 
  geom_bar( stat ="identity", position ="identity", colour ="black", size = 0.25) + 
  scale_fill_manual( values = c("#CCEEFF", "#FFDDDD"), guide = FALSE)
```

## 6. Adjusting Bar Width and Spacing
To make the bars narrower or wider, set **width** in geom_bar(). The default value is 0.9. 
```{r}
## default 
ggplot( pg_mean, aes( x = group, y = weight)) + geom_bar( stat ="identity")

## for narrower bars
ggplot( pg_mean, aes( x = group, y = weight)) + geom_bar( stat ="identity", width = 0.5)

## for wider bars
ggplot( pg_mean, aes( x = group, y = weight)) + geom_bar( stat ="identity", width = 1) 

## To add space between bars within a group, make width smaller and set the value for position_dodge
ggplot( cabbage_exp, aes( x = Date, y = Weight, fill = Cultivar)) + 
  geom_bar( stat ="identity", width = 0.5, position ="dodge")

## And with some space between the bars:
ggplot( cabbage_exp, aes( x = Date, y = Weight, fill = Cultivar)) + 
  geom_bar( stat ="identity", width = 0.5, position = position_dodge( 0.7))
```

## 7. Making a Stacked Bar Graph
```{r}
ggplot( cabbage_exp, aes( x = Date, y = Weight, fill = Cultivar)) + 
  geom_bar( stat ="identity")

## reverse the order of items in the legend by using guides()
ggplot( cabbage_exp, aes( x = Date, y = Weight, fill = Cultivar)) + 
  geom_bar( stat ="identity") + 
  guides( fill = guide_legend( reverse = TRUE))

## reverse the stacking order
ggplot( cabbage_exp, aes( x = Date, y = Weight, fill = Cultivar, order = desc( Cultivar))) + 
  geom_bar( stat ="identity")

## scale_fill_brewer 
ggplot( cabbage_exp, aes( x = Date, y = Weight, fill = Cultivar)) + 
  geom_bar( stat = "identity", colour = "black") + 
  guides( fill = guide_legend( reverse = TRUE)) + 
  scale_fill_brewer( palette = "Pastel1")
```

## 8. Making a Proportional Stacked Bar Graph
```{r}
ce <- ddply( cabbage_exp, "Date", transform, percent_weight = Weight / sum( Weight) * 100)

ggplot( ce, aes( x = Date, y = percent_weight, fill = Cultivar)) + 
  geom_bar( stat ="identity")

ce <- ddply( cabbage_exp, "Date", transform, percent_weight = Weight / sum( Weight) * 100)

ggplot( ce, aes( x = Date, y = percent_weight, fill = Cultivar)) + 
  geom_bar( stat ="identity", colour ="black") + 
  guides( fill = guide_legend( reverse = TRUE)) + 
  scale_fill_brewer( palette ="Pastel1")
```

## 9. MAdding Labels to a Bar Graph
```{r}
# Below the top 
ggplot( cabbage_exp, aes( x = interaction( Date, Cultivar), y = Weight)) + 
  geom_bar( stat ="identity") + 
  geom_text( aes( label = Weight), vjust = 1.5, colour ="white") 

# Above the top 
ggplot( cabbage_exp, aes( x = interaction( Date, Cultivar), y = Weight)) + 
  geom_bar( stat ="identity") + 
  geom_text( aes( label = Weight), vjust =-0.2)

# Adjust y limits to be a little higher 
ggplot( cabbage_exp, aes( x = interaction( Date, Cultivar), y = Weight)) + 
  geom_bar( stat ="identity") + geom_text( aes( label = Weight), vjust =-0.2) + 
  ylim( 0, max( cabbage_exp$Weight) * 1.05) 

# Map y positions slightly above bar top - y range of plot will auto-adjust 
ggplot( cabbage_exp, aes( x = interaction( Date, Cultivar), y = Weight)) + 
  geom_bar( stat ="identity") + geom_text( aes( y = Weight + 0.1, label = Weight))

# Sort by the day and sex columns 
ce <- arrange( cabbage_exp, Date, Cultivar)

## Get the cumulative sum
ce <- ddply( ce, "Date", transform, label_y = cumsum( Weight))

ggplot( ce, aes( x = Date, y = Weight, fill = Cultivar)) + 
  geom_bar( stat ="identity") + 
  geom_text( aes( y = label_y, label = Weight), vjust = 1.5, colour ="white")

# Calculate y position, placing it in the middle 
ce <- ddply( ce, "Date", transform, label_y = cumsum( Weight)-0.5* Weight) 
ggplot( ce, aes( x = Date, y = Weight, fill = Cultivar)) + 
  geom_bar( stat ="identity") + 
  geom_text( aes( y = label_y, label = Weight), colour ="white")

## add labels in the middle with a smaller font using size, add a “kg”
ggplot( ce, aes( x = Date, y = Weight, fill = Cultivar)) + 
  geom_bar( stat ="identity", colour ="black") + 
  geom_text( aes( y = label_y, label = paste( format( Weight, nsmall = 2), "kg")), size = 4) + 
  guides( fill = guide_legend( reverse = TRUE)) + scale_fill_brewer( palette ="Pastel1")
```

## Making a Cleveland Dot Plot
```{r}
# Take the top 25 from the tophitters data set 
tophit <- tophitters2001[ 1: 25, ] 
ggplot( tophit, aes( x = avg, y = name)) + geom_point()

## reorder( name, avg),
ggplot( tophit, aes( x = avg, y = reorder( name, avg))) + 
  geom_point( size = 3) + 
  theme_bw() + 
  theme( panel.grid.major.x = element_blank(), 
         panel.grid.minor.x = element_blank(), 
         panel.grid.major.y = element_line( colour ="grey60", linetype ="dashed"))

## swap the axes
ggplot( tophit, aes( x = reorder( name, avg), y = avg)) + 
  geom_point( size = 3) + 
  theme_bw() + 
  theme( axis.text.x = element_text( angle = 60, hjust = 1), 
         panel.grid.major.y = element_blank(), 
         panel.grid.minor.y = element_blank(), 
         panel.grid.major.x = element_line( colour ="grey60", linetype ="dashed"))

# Get the names, sorted first by lg, then by avg 
nameorder <- tophit$name[ order( tophit$lg, tophit$avg)] 

# Turn name into a factor, with levels in the order of nameorder 
tophit$name <- factor( tophit$name, levels = nameorder)

ggplot( tophit, aes( x = avg, y = name)) + 
  geom_segment( aes( yend = name), xend = 0, colour ="grey50") + 
  geom_point( size = 3, aes( colour = lg)) + scale_colour_brewer( palette ="Set1", limits = c("NL","AL")) + 
  theme_bw() + 
  theme( panel.grid.major.y = element_blank(), # No horizontal grid lines 
         legend.position = c( 1, 0.55), # Put legend inside plot area 
         legend.justification = c( 1, 0.5))

## Another way to separate the two groups is to use facets,
ggplot( tophit, aes( x = avg, y = name)) + 
  geom_segment( aes( yend = name), xend = 0, colour ="grey50") + 
  geom_point( size = 3, aes( colour = lg)) + 
  scale_colour_brewer( palette ="Set1", limits = c("NL","AL"), guide = FALSE) + 
  theme_bw() + 
  theme( panel.grid.major.y = element_blank()) + 
  facet_grid( lg ~ ., scales ="free_y", space ="free_y")
```



