---
title: "5_Scatter_Plots"
author: "Gino Tesei"
date: "December 12, 2014"
output: html_document
---

## 1. Making a Basic Scatter Plot 
```{r}
library(ggplot2)
library(gcookbook) # For the data set 

library(plyr) ##

# List the two columns we'll use 
head(heightweight[, c("ageYear", "heightIn")])
ggplot( heightweight, aes( x = ageYear, y = heightIn)) + 
  geom_point()
 
## To use different shapes in a scatter plot, set shape.
ggplot( heightweight, aes( x = ageYear, y = heightIn)) + geom_point( shape = 21)

## The size of the points can be controlled with size.
ggplot( heightweight, aes( x = ageYear, y = heightIn)) + geom_point( size = 1.5)
```

## 2. Grouping Data Points by a Variable Using Shape or Color
```{r}
# Show the three columns we'll use 
head(heightweight[, c("sex", "ageYear", "heightIn")])

ggplot( heightweight, aes( x = ageYear, y = heightIn, colour = sex)) + 
  geom_point() 

ggplot( heightweight, aes( x = ageYear, y = heightIn, shape = sex)) + 
  geom_point() 

## It is possible to map a variable to both shape and colour,
ggplot( heightweight, aes( x = ageYear, y = heightIn, shape = sex, colour = sex)) + geom_point()

## set different shapes and colors for the grouping variables
ggplot( heightweight, aes( x = ageYear, y = heightIn, shape = sex, colour = sex)) + 
  geom_point() + 
  scale_shape_manual( values = c(1,2)) + 
  scale_colour_brewer( palette ="Set1")
```

## 3. Using Different Point Shapes
```{r}
ggplot( heightweight, aes( x = ageYear, y = heightIn)) + 
  geom_point( shape = 3)

# Use slightly larger points and use a shape scale with custom values 
ggplot( heightweight, aes( x = ageYear, y = heightIn, shape = sex)) + 
  geom_point( size = 3) + 
  scale_shape_manual( values = c( 1, 4))

## It’s possible to have the shape represent one variable and the fill (empty or solid) represent another variable.
# Make a copy of the data 
hw <- heightweight 

# Categorize into < 100 and > = 100 groups 
hw$weightGroup <- cut( hw$weightLb, breaks = c(-Inf, 100, Inf), 
                       labels = c("< 100", "> = 100")) 

# Use shapes with fill and color, and use colors that are empty (NA) and # filled 
ggplot( hw, aes( x = ageYear, y = heightIn, shape = sex, fill = weightGroup)) + 
  geom_point( size = 2.5) + scale_shape_manual( values = c( 21, 24)) + 
  scale_fill_manual( values = c( NA, "black"), guide = guide_legend( override.aes = list( shape = 21)))
```


## 4. Mapping a Continuous Variable to Color or Size
```{r}
# List the four columns we'll use 
head(heightweight[, c("sex", "ageYear", "heightIn", "weightLb")])

ggplot( heightweight, aes( x = ageYear, y = heightIn, colour = weightLb)) + 
  geom_point() 

ggplot( heightweight, aes( x = ageYear, y = heightIn, size = weightLb )) + 
  geom_point() 

## set the fill gradient to go from black to white and make the points larger
ggplot( heightweight, aes( x = ageYear, y = heightIn, fill = weightLb)) + 
  geom_point( shape = 21, size = 2.5) + 
  scale_fill_gradient( low ="black", high ="white") 

# Using guide_legend() will result in a discrete legend instead of a colorbar 
ggplot( heightweight, aes( x = ageYear, y = heightIn, fill = weightLb)) + 
  geom_point( shape = 21, size = 2.5) + 
  scale_fill_gradient( low ="black", high ="white", breaks = seq( 70, 170, by = 20), guide = guide_legend())

## best 
ggplot( heightweight, aes( x = ageYear, y = heightIn, size = weightLb, colour = sex)) + 
  geom_point( alpha =.5) + 
  scale_size_area() + # Make area proportional to numeric value 
  scale_colour_brewer( palette ="Set1")
```

## 5. Dealing with Overplotting
You have many points and they obscure each other.
```{r}
sp <- ggplot( diamonds, aes( x = carat, y = price)) 
sp + 
  geom_point()

## We can make the points semitransparent using alpha,
sp + 
  geom_point( alpha =.1) 

sp + 
  geom_point( alpha =.01)

## Another solution is to bin the points into rectangles and map the density of the points to the fill color of the rectangles,
sp + 
  stat_bin2d() 

sp + 
  stat_bin2d( bins = 50) + 
  scale_fill_gradient( low ="lightblue", high ="red", limits = c( 0, 6000))
```

Another alternative is to bin the data into hexagons instead of rectangles, with stat_binhex. It works just like stat_bin2d().
```{r}
library( hexbin) 
sp + 
  stat_binhex() + 
  scale_fill_gradient( low ="lightblue", high ="red", limits = c( 0, 8000))

sp + 
  stat_binhex() + 
  scale_fill_gradient( low ="lightblue", high ="red", breaks = c( 0, 250, 500, 1000, 2000, 4000, 6000), 
                       limits = c( 0, 6000))
```

Overplotting can also occur when the data is discrete on one or both axes. In these cases, you can randomly jitter the points with position_jitter().

```{r}
sp1 <- ggplot( ChickWeight, aes( x = Time, y = weight)) 

sp1 + geom_point() 

sp1 + geom_point( position ="jitter") 

# Could also use geom_jitter(), which is equivalent 
sp1 + geom_point( position = position_jitter( width =.5, height = 0))

## 
sp1 + geom_boxplot( aes( group = Time))
```

## 6. Adding Fitted Regression Model Lines
```{r}
# The base plot 
sp <- ggplot( heightweight, aes( x = ageYear, y = heightIn)) 

# By default, stat_smooth() also adds a 95% confidence region for the regression fit
sp + 
  geom_point() + 
  stat_smooth( method = lm)

# 99% confidence region 
sp + 
  geom_point() + 
  stat_smooth( method = lm, level = 0.99)

# No confidence region 
sp + 
  geom_point() + 
  stat_smooth( method = lm, se = FALSE)

# The default color of the fit line is blue. This can be change by setting colour.
sp + 
  geom_point( colour ="grey60") + 
  stat_smooth( method = lm, se = FALSE, colour ="black")

# If you add stat_smooth() without specifying the method, it will use a loess (locally weighted polynomial)
sp + 
  geom_point( colour ="grey60") + 
  stat_smooth() 

sp + 
  geom_point( colour ="grey60") + 
  stat_smooth( method = loess)
```
 
Another common type of model fit is a logistic regression.
```{r}
library( MASS) # For the data set 
b <- biopsy 
b$classn[ b$class =="benign"] <- 0 
b$classn[ b$class =="malignant"] <- 1 
head(b)

# we’ll just look at the relationship of V1 (clump thickness) and the class of the tumor.
ggplot( b, aes( x = V1, y = classn)) + 
  geom_point( position = position_jitter( width = 0.3, height = 0.06), alpha = 0.4, shape = 21, size = 1.5) + 
  stat_smooth( method = glm, family = binomial)

# If your scatter plot has points grouped by a factor, using colour or shape, one fit line will be drawn for each group.
sps <- ggplot( heightweight, aes( x = ageYear, y = heightIn, colour = sex)) + 
  geom_point() + 
  scale_colour_brewer( palette ="Set1") 
sps + geom_smooth()

# If you want the lines to extrapolate from the data 
sps + 
  geom_smooth( method = lm, se = FALSE, fullrange = TRUE)
```

## 7. If you want the lines to extrapolate from the data
You have already created a fitted regression model object for a data set, and you want to plot the lines for that model.
```{r}
model <- lm( heightIn ~ ageYear + I(ageYear^2), heightweight)  
model

# Create a data frame with ageYear column, interpolating across range 
xmin <- min( heightweight$ageYear) 
xmax <- max( heightweight$ageYear) 
predicted <- data.frame( ageYear = seq( xmin, xmax, length.out = 100))
predicted$heightIn <- predict( model, predicted) 
head(predicted)

# We can now plot the data points along with the values predicted from the model
sp <- ggplot( heightweight, aes( x = ageYear, y = heightIn)) + 
  geom_point( colour ="grey40") 

sp + geom_line( data = predicted, size = 1)

# Given a model, predict values of yvar from xvar 
# This supports one predictor and one predicted variable
# xrange: If NULL, determine the x range from the model object. If a vector with 
# two numbers, use those as the min and max of the prediction range. 
# samples: Number of samples across the x range. 
# ...: Further arguments to be passed to predict() 
predictvals <- function( model, xvar, yvar, xrange = NULL, samples = 100, ...) { 
  # If xrange isn't passed in, determine xrange from the models. 
  # Different ways of extracting the x range, depending on model type 
  if (is.null(xrange)) { 
    if (any( class( model) %in% c("lm", "glm"))) 
      xrange <- range( model $ model[[ xvar]]) 
    else if (any( class( model) %in% "loess")) xrange <- range( model $ x) 
    } 
  newdata <- data.frame( x = seq( xrange[1], xrange[2], length.out = samples)) 
  names( newdata) <- xvar 
  newdata[[ yvar]] <- predict( model, newdata = newdata, ...) 
  newdata 
  }

modlinear <- lm( heightIn ~ ageYear, heightweight) 
modloess <- loess( heightIn ~ ageYear, heightweight)

lm_predicted <- predictvals( modlinear, "ageYear", "heightIn") 
loess_predicted <- predictvals( modloess, "ageYear", "heightIn") 

sp + 
  geom_line( data = lm_predicted, colour ="red", size =.8) + 
  geom_line( data = loess_predicted, colour ="blue", size =.8)
```

For glm models that use a nonlinear link function, you need to specify type =" response"

```{r}
library( MASS) # For the data set 
b <- biopsy 
b$classn[ b$class =="benign"] <- 0 
b$classn[ b$class =="malignant"] <- 1 
head(b)

# glm model 
fitlogistic <- glm( classn ~ V1, b, family = binomial)

# Get predicted values 
glm_predicted <- predictvals( fitlogistic, "V1", "classn", type ="response") 

ggplot( b, aes( x = V1, y = classn)) + 
  geom_point( position = position_jitter( width =.3, height =.08), alpha = 0.4, shape = 21, size = 1.5) + 
  geom_line( data = glm_predicted, colour ="#1177FF", size = 1)
```

## 8. Adding Fitted Lines from Multiple Existing Models
With the heightweight data set, we’ll make a linear model with lm() for each of the levels of sex, and put those model objects in a list. The model building is done with a function, make_model() 
```{r}
make_model <- function( data) { 
  lm( heightIn ~ ageYear, data) 
  }
models <- dlply( heightweight, "sex", .fun = make_model) # Print out the list of two lm objects, f and m models
models 

# Now that we have the list of model objects, we can run predictvals() to get predicted values from each model, using the ldply() function
predvals <- ldply( models, .fun = predictvals, xvar ="ageYear", yvar ="heightIn") 
head(predvals)

# Finally, we can plot the data with the predicted values
ggplot( heightweight, aes( x = ageYear, y = heightIn, colour = sex)) + 
  geom_point() + 
  geom_line( data = predvals)

predvals <- ldply( models, .fun = predictvals, xvar ="ageYear", yvar ="heightIn", 
                   xrange = range( heightweight$ageYear))
ggplot( heightweight, aes( x = ageYear, y = heightIn, colour = sex)) + 
  geom_point() + geom_line( data = predvals)
```

## 9. Adding Annotations with Model Coefficients
```{r}
model <- lm( heightIn ~ ageYear, heightweight) 
summary( model)
 
# First generate prediction data 
pred <- predictvals( model, "ageYear", "heightIn") 
sp <- ggplot( heightweight, aes( x = ageYear, y = heightIn)) + 
  geom_point() + geom_line( data = pred) 

##
sp + 
  annotate("text", label ="r^2 = 0.42", x = 16.5, y = 52)

##
sp + 
  annotate("text", label ="r^2 == 0.42", parse = TRUE, x = 16.5, y = 52)

## It’s possible to automatically extract values from the model object and build an expression
eqn <- as.character( 
  as.expression( substitute( italic( y) == a + b * italic( x) * "," ~~ italic(r)^2 ~ "=" ~ r2, 
                             list( a = format( coef( model)[1], digits = 3), 
                                   b = format( coef( model)[2], digits = 3), 
                                   r2 = format( summary(model)$r.squared, 
                                                digits = 2) )))) 
eqn

# Parsing turns it into an expression
parse( text = eqn)

# Now that we have the expression string, we can add it to the plot.
sp + annotate("text", label = eqn, parse = TRUE, x = Inf, y =-Inf, hjust = 1.1, vjust =-.5)
```

## 10. Adding Marginal Rugs to a Scatter Plot
```{r}
ggplot( faithful, aes( x = eruptions, y = waiting)) + 
  geom_point() + 
  geom_rug()

# To reduce the overplotting, we can jitter the line positions and make them slightly thinner by specifying size
ggplot( faithful, aes( x = eruptions, y = waiting)) + 
  geom_point() + 
  geom_rug( position ="jitter", size =.2)
```

## 11. Labeling Points in a Scatter Plot
```{r}
# we’ll just take the subset of countries that spent more than $ 2000 USD per capita:
subset( countries, Year == 2009 & healthexp > 2000)

# To manually add annotations, use annotate(),
sp <- ggplot( subset( countries, Year == 2009 & healthexp > 2000), aes( x = healthexp, y = infmortality)) + 
  geom_point() 

sp + 
  annotate("text", x = 4350, y = 5.4, label ="Canada") + 
  annotate("text", x = 7400, y = 6.8, label ="USA")

# To automatically add the labels from your data we use geom_text() and map a column that is a factor or character vector to the label aesthetic.
sp + 
  geom_text( aes( label = Name), size = 4)

sp + 
  geom_text( aes( label = Name), size = 4, vjust = 0)

# Add a little extra to y 
sp + 
  geom_text( aes( y = infmortality +.1, label = Name), size = 4, vjust = 0)

sp + 
  geom_text( aes( label = Name), size = 4, hjust = 0) 

sp + 
  geom_text( aes( x = healthexp + 100, label = Name), size = 4, hjust = 0)

# you can add a new column to your data frame containing just the labels you want.
cdat <- subset( countries, Year == 2009 & healthexp > 2000) 
cdat$Name1 <- cdat$Name
idx <- cdat$Name1 %in% c(" Canada", "Ireland", "United Kingdom", "United States", "New Zealand", "Iceland", "Japan", "Luxembourg", "Netherlands", "Switzerland") 
head(idx)

# Then we’ll use that Boolean vector to overwrite all the other entries in Name1 with NA: 
cdat$Name1[!idx] <- NA

# Now we can make the plot
ggplot( cdat, aes( x = healthexp, y = infmortality)) + 
  geom_point() + 
  geom_text( aes( x = healthexp + 100, label = Name1), size = 4, hjust = 0) + 
  xlim( 2000, 10000)
```

## 12. Creating a Balloon Plot
```{r}
p <- ggplot( cdat, aes( x = healthexp, y = infmortality, size = GDP)) + 
  geom_point( shape = 21, colour ="black", fill ="cornsilk") 

# GDP mapped to radius (default with scale_size_continuous) 
p 

# GDP mapped to area instead, and larger circles 
p + scale_size_area( max_size = 15)

# Add up counts for male and female 
hec <- HairEyeColor[,,"Male"] + HairEyeColor[,,"Female"] 

# Convert to long format 
library( reshape2) 
hec <- melt( hec, value.name ="count") 

ggplot( hec, aes( x = Eye, y = Hair)) + 
  geom_point( aes( size = count), shape = 21, colour ="black", fill ="cornsilk") + 
  scale_size_area( max_size = 20, guide = FALSE) + 
  geom_text( aes( y = as.numeric(Hair)-sqrt(count)/ 22, label = count), vjust = 1, colour ="grey60", size = 4)
```

## 13. Making a Scatter Plot Matrix
```{r}
c2009 <- subset( countries, Year == 2009, select = c( Name, GDP, laborrate, healthexp, infmortality)) 
head(c2009)

# make the scatter plot matrix
pairs( c2009[, 2: 5])

panel.cor <- function( x, y, digits = 2, prefix ="", cex.cor, ...) { 
  usr <- par("usr") 
  on.exit( par(usr)) 
  par( usr = c( 0, 1, 0, 1)) 
  r <- abs( cor( x, y, use ="complete.obs")) 
  txt <- format( c( r, 0.123456789), digits = digits)[1] 
  txt <- paste( prefix, txt, sep ="") 
  if( missing( cex.cor)) cex.cor <- 0.8 / strwidth(txt) 
  text( 0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2) 
  }

panel.hist <- function( x, ...) { 
  usr <- par("usr") 
  on.exit( par(usr)) 
  par( usr = c( usr[1:2], 0, 1.5) ) 
  h <- hist( x, plot = FALSE) 
  breaks <- h$breaks 
  nB <- length(breaks) 
  y <- h$counts 
  y <- y/max( y) 
  rect( breaks[-nB], 0, breaks[-1], y, col ="white", ...) 
  }

pairs( c2009[,2:5], upper.panel = panel.cor, 
       diag.panel = panel.hist, 
       lower.panel = panel.smooth)

# It may be more desirable to use linear regression lines instead of LOWESS lines.
panel.lm <- function (x, y, col = par("col"), bg = NA, pch = par("pch"), cex = 1, col.smooth = "black", ...) { 
  points( x, y, pch = pch, col = col, bg = bg, cex = cex) 
  abline( stats::lm( y ~ x), col = col.smooth, ...) 
  }

pairs( c2009[, 2:5], pch =".", 
       upper.panel = panel.cor, 
       diag.panel = panel.hist, 
       lower.panel = panel.lm)
```
