---
title: "8_Axes"
author: "Gino Tesei"
date: "December 13, 2015"
output: html_document
---

## 1. Swapping X- and Y-Axes
```{r}
library(ggplot2)
library(gcookbook) # For the data set 

library(plyr)

ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() 

## Use coord_flip() to flip the axes
ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() + coord_flip()

# If the x variable is a factor, the order can be reversed by using scale_x_discrete() with limits = rev( levels(...))
ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() + 
  coord_flip() + 
  scale_x_discrete( limits = rev( levels( PlantGrowth$group)))
```

## 2. Setting the Range of a Continuous Axis
```{r}
p <- ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() 

# Display the basic graph 
p 

# Display the graph with longet y axis  
p + ylim( 0, max( PlantGrowth$weight))
```

## 3. Reversing a Continuous Axis
```{r}
ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() + 
  scale_y_reverse() 

# Similar effect by specifying limits in reversed order 
ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() + ylim( 6.5, 3.5)

# Like scale_y_continuous(), scale_y_reverse() does not work with ylim. 
# The same is true for the x-axis properties. 
# If you want to reverse an axis and set its range, you must do it within the scale_y_reverse() statement, by setting the limits in reversed order
ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() + 
  scale_y_reverse( limits = c( 8, 0))
```

## 4. Changing the Order of Items on a Categorical Axis
```{r}
# To manually set the order of items on the axis, specify limits with a vector of the levels in the desired order.
p <- ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() 

p + 
  scale_x_discrete( limits = c("ctrl","trt1"))

# To reverse the order, set limits = rev( levels(...)),
p + 
  scale_x_discrete( limits = rev( levels( PlantGrowth$group)))
```

## 5. Setting the Scaling Ratio of the X- and Y-Axes
```{r}
sp <- ggplot( marathon, aes( x = Half, y = Full)) + 
  geom_point() 

# Use coord_fixed(). This will result in a 1:1 scaling between the x- and y-axes
sp + 
  coord_fixed()

# It’s also helpful to set the tick spacing to be the same, by setting breaks in scale_y_continuous() and scale_x_continuous()
sp + 
  coord_fixed() + 
  scale_y_continuous( breaks = seq( 0, 420, 30)) + 
  scale_x_continuous( breaks = seq( 0, 420, 30))

# If, instead of an equal ratio, you want some other fixed ratio between the axes, set the ratio parameter.
sp + 
  coord_fixed( ratio = 1/2) + 
  scale_y_continuous( breaks = seq( 0, 420, 30)) + 
  scale_x_continuous( breaks = seq( 0, 420, 15))
```

## 6. Setting the Positions of Tick Marks
```{r}
# Usually ggplot() does a good job of deciding where to put the tick marks
ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() 

# but if you want to change them, set breaks in the scale
ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() + 
  scale_y_continuous( breaks = c( 4, 4.25, 4.5, 5, 6, 8))

# Discrete axis: setting breaks will change which of the levels are labeled, but will not remove them or change their order
ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() + 
  scale_x_discrete( limits = c("trt2", "ctrl"), breaks ="ctrl")
```

## 7. Removing Tick Marks and Labels
```{r}

p <- ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() 

# default 
p

# To remove just the tick labels use theme( axis.text.y = element_blank())
p + 
  theme( axis.text.y = element_blank())

# To remove the tick marks, use theme( axis.ticks = element_blank()).
p + 
  theme( axis.ticks = element_blank(), axis.text.y = element_blank())

# To remove the tick marks, the labels, and the grid lines, set breaks to NULL
p + 
  scale_y_continuous( breaks = NULL)
```

## 8. Changing the Text of Tick Labels
```{r}
hwp <- ggplot( heightweight, aes( x = ageYear, y = heightIn)) + 
  geom_point() 

# default 
hwp

# To set arbitrary labels pass values to breaks and labels in the scale 
hwp + 
  scale_y_continuous( breaks = c( 50, 56, 60, 66, 72), labels = c("Tiny", "Really\nshort", "Short", "Medium", "Tallish"))

# we can define a formatter function, which takes in a value and returns the corresponding string.
footinch_formatter <- function( x) { 
  foot <- floor( x/12) 
  inch <- x %% 12 
  return( paste( foot, "'", inch, "\"", sep ="")) 
  }

# we can pass our function to the scale, using the labels parameter
hwp + 
  scale_y_continuous( labels = footinch_formatter)

# We can instead have ggplot() set tick marks every four inches, by specifying breaks
hwp + 
  scale_y_continuous( breaks = seq( 48, 72, 4), labels = footinch_formatter)

# Another common task is to convert time measurements to HH:MM:SS format, or something similar.
timeHMS_formatter <- function( x) { 
  h <- floor( x/ 60)
  m <- floor( x %% 60) 
  s <- round( 60*( x %% 1)) 
  
  # Round to nearest second 
  lab <- sprintf("%02d:%02d:%02d", h, m, s) 
  
  # Format the strings as HH:MM:SS 
  lab <- gsub("^00:", "", lab) 
  
  # Remove leading 00: if present 
  lab <- gsub("^0", "", lab) 
  
  # Remove leading 0 if present 
  
  return(lab) 
  }

timeHMS_formatter( c(.33, 50, 51.25, 59.32, 60, 60.1, 130.23))
```

## 9. Changing the Appearance of Tick Labels
```{r}
# manually set the labels to be long — long enough that they overlap:
bp <- ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() + scale_x_discrete( breaks = c("ctrl", "trt1", "trt2"), 
                                     labels = c("Control", "Treatment 1", "Treatment 2")) 
bp

# To rotate the text 90 degrees counterclockwise
bp + 
  theme( axis.text.x = element_text( angle = 90, hjust = 1, vjust =.5))

# Rotating the text 30 degrees
bp + 
  theme( axis.text.x = element_text( angle = 30, hjust = 1, vjust = 1))

# Besides rotation, other text properties, such as size, style (bold/ italic/ normal), and the font family (such as Times or Helvetica) can be set with element_text()
bp + 
  theme( axis.text.x = element_text( family ="Times", face ="italic", colour ="darkred", size = rel( 0.9)))
```

## 10. Changing the Text of Axis Labels
```{r}
hwp <- ggplot( heightweight, aes( x = ageYear, y = heightIn, colour = sex)) + 
  geom_point() 

# With default axis labels 
hwp

# Set the axis labels 
hwp + 
  xlab("Age\nin years") + 
  ylab("Height in inches")
```

## 11. Removing Axis Labels
```{r}
p <- ggplot( PlantGrowth, aes( x = group, y = weight)) + 
  geom_boxplot() 

# We’ll hide the x-axis
p + 
  theme( axis.title.x = element_blank())
```

## 12. Changing the Appearance of Axis Labels
```{r}
# To change the appearance of the x-axis use axis.title.x
hwp <- ggplot( heightweight, aes( x = ageYear, y = heightIn)) + 
  geom_point() 

hwp + 
  theme( axis.title.x = element_text( face ="italic", colour ="darkred", size = 14))

# y axis 
hwp + 
  ylab(" Height\n( inches)") + 
  theme( axis.title.y = element_text( angle = 90, face ="italic", colour ="darkred", size = 14))
```

## 13. Showing Lines Along the Axes
```{r}
# You want to display lines along the x- and y-axes, but not on the other sides of the graph.
p <- ggplot( heightweight, aes( x = ageYear, y = heightIn)) + 
  geom_point() 

p + theme( axis.line = element_line( colour ="black"))

# With thick lines, only half overlaps 
p + 
  theme_bw() + 
  theme( panel.border = element_blank(), axis.line = element_line( colour ="black", size = 4)) 

# Full overlap 
p + 
  theme_bw() + 
  theme( panel.border = element_blank(), axis.line = element_line( colour ="black", size = 4, lineend ="square"))
```

## 14. Using a Logarithmic Axis
```{r}
library( MASS)

# The base plot 
p <- ggplot( Animals, aes( x = body, y = brain, label = rownames( Animals))) + 
  geom_text( size = 3) 

p 

# With logarithmic x and y scales 
p + 
  scale_x_log10() + 
  scale_y_log10()

# breaks 
p + 
  scale_x_log10( breaks = 10^(-1:5)) + scale_y_log10( breaks = 10^(0:3))

# Another way to use log axes is to transform the data before mapping it to the x and y coordinates
ggplot( Animals, aes( x = log10( body), y = log10( brain), label = rownames( Animals))) + 
  geom_text( size = 3)
```

## 15. Adding Ticks for a Logarithmic Axis
```{r}
# Use annotation_logticks()
library( MASS) # For the data set 
library( scales) # For the trans and format functions 

ggplot( Animals, aes( x = body, y = brain, label = rownames( Animals))) + 
  geom_text( size = 3) + annotation_logticks() + 
  scale_x_log10( breaks = trans_breaks("log10", function(x) 10^x), labels = trans_format("log10", math_format(10^.x))) +
  scale_y_log10( breaks = trans_breaks("log10", function(x) 10^x), labels = trans_format("log10", math_format( 10^.x)))

# To get the colors of the tick marks and the grid lines to match up a bit better, you can use theme_bw().
ggplot( Animals, aes( x = body, y = brain, label = rownames( Animals))) + 
  geom_text( size = 3) + 
  annotation_logticks() + 
  scale_x_log10( breaks = trans_breaks("log10", function(x) 10^x), 
                 labels = trans_format("log10", math_format(10^.x)), 
                 minor_breaks = log10( 5) + -2: 5) + 
  scale_y_log10( breaks = trans_breaks("log10", function(x) 10^x), 
                 labels = trans_format("log10", math_format(10^.x)), 
                 minor_breaks = log10(5) + -1:3) + 
  coord_fixed() + 
  theme_bw()
```

## 15. Making a Circular Graph
```{r}
ggplot( wind, aes( x = DirCat, fill = SpeedCat)) + 
  geom_histogram( binwidth = 15, origin = -7.5) + 
  coord_polar() + 
  scale_x_continuous( limits = c( 0,360))
 
# we can make the plot a little prettier by reversing the legend, using a different palette, adding an outline, and setting the breaks to some more familiar numbers
ggplot( wind, aes( x = DirCat, fill = SpeedCat)) + 
  geom_histogram( binwidth = 15, origin =-7.5, colour ="black", size =.25) + 
  guides( fill = guide_legend( reverse = TRUE)) + 
  coord_polar() + 
  scale_x_continuous( limits = c( 0,360), breaks = seq( 0, 360, by = 45), minor_breaks = seq( 0, 360, by = 15)) + 
  scale_fill_brewer()

# Put mdeaths time series data into a data frame 
md <- data.frame( deaths = as.numeric( mdeaths), 
                  month = as.numeric( cycle( mdeaths))) 

# Calculate average number of deaths in each month 
library( plyr) # For the ddply() function 
md <- ddply( md, "month", summarise, deaths = mean( deaths)) 
str(md)

# Make the base plot 
p <- ggplot( md, aes( x = month, y = deaths)) + 
  geom_line() + scale_x_continuous( breaks = 1:12) 

# With coord_polar 
p + 
  coord_polar()

# With coord_polar and y (r) limits going to zero 
p + 
  coord_polar() + ylim( 0, max( md$deaths))

# also xlim 
p + 
  coord_polar() + 
  ylim( 0, max( md$deaths)) + 
  xlim( 0, 12)

# Connect the lines by adding a value for 0 that is the same as 12 
mdx <- md[ md$month == 12, ] 
mdx $ month <- 0 
mdnew <- rbind( mdx, md) 

# Make the same plot as before, but with the new data, by using %+% 
p %+% mdnew + 
  coord_polar() + 
  ylim( 0, max( md$deaths))
```

## 15. Using Dates on an Axis
```{r}
# Look at the structure 
str( economics)

ggplot( economics, aes( x = date, y = psavert)) + 
  geom_line()

# Take a subset of economics 
econ <- subset( economics, date >= as.Date("1992-05-01") & date < as.Date("1993-06-01")) 

# Base plot - without specifying breaks 
p <- ggplot( econ, aes( x = date, y = psavert)) + 
  geom_line() 

p


# Specify breaks as a Date vector 
datebreaks <- seq( as.Date("1992-06-01"), as.Date("1993-06-01"), by ="2 month") 

# Use breaks, and rotate text labels 
p + 
  scale_x_date( breaks = datebreaks) + 
  theme( axis.text.x = element_text( angle = 30, hjust = 1))

# Line graph with date format specified
library( scales) 
p + 
  scale_x_date( breaks = datebreaks, labels = date_format("% Y %b")) + 
  theme( axis.text.x = element_text( angle = 30, hjust = 1))
```

## 16. Using Relative Times on an Axis
```{r}
# Convert WWWusage time-series object to data frame 
www <- data.frame( minute = as.numeric( time( WWWusage)), users = as.numeric( WWWusage)) 
# Define a formatter function - converts time in minutes to a string 
timeHM_formatter <- function( x) { 
  h <- floor( x/ 60) 
  m <- floor( x %% 60) 
  lab <- sprintf("%d:%02d", h, m) 
  # Format the strings as HH:MM 
  return( lab) 
  } 

# Default x axis 
ggplot( www, aes( x = minute, y = users)) + 
  geom_line() 

# With formatted times 
ggplot( www, aes( x = minute, y = users)) + 
  geom_line() + 
  scale_x_continuous( name =" time", breaks = seq( 0, 100, by = 10), labels = timeHM_formatter)

# To convert to HH:MM:SS format, you can use the following formatter function
timeHMS_formatter <- function(x) { 
  h <- floor( x/ 3600) 
  m <- floor(( x/ 60) %% 60) 
  s <- round( x %% 60) 
  
  # Round to nearest second 
  lab <- sprintf("%02d:%02d:%02d", h, m, s) 
  # Format the strings as HH:MM:SS 
  lab <- sub("^00:", "", lab) 
  # Remove leading 00: if present 
  lab <- sub("^0", "", lab) 
  # Remove leading 0 if present 
  return( lab)
  }

ggplot( www, aes( x = minute, y = users)) + 
  geom_line() +
  scale_x_continuous( breaks = c( 0, 20, 40, 60, 80, 100), labels = c("0:00", "0:20", "0:40", "1:00", "1:20", "1:40"))
```





