---
title: "Exploratory_2_ip_url"
author: "Gino Tesei"
date: "May 25, 2015"
output: html_document
---


```{r, echo=FALSE , results="hide" , include=FALSE}
library(caret)
library(Hmisc)
library(data.table)
library(kernlab)
library(subselect)
library(plyr)
library(binhf)
library(fBasics)
library(lattice)

getBasePath = function (type = "data") {
  ret = ""
  base.path1 = ""
  base.path2 = ""
  
  if(type == "data") {
    base.path1 = "C:/docs/ff/gitHub/fast-furious/dataset/facebook-recruiting-iv-human-or-bot"
    base.path2 = "/Users/gino/kaggle/fast-furious/gitHub/fast-furious/dataset/facebook-recruiting-iv-human-or-bot/"
  } else if (type == "code") {
    base.path1 = "C:/docs/ff/gitHub/fast-furious/competitions/facebook-recruiting-iv-human-or-bot"
    base.path2 = "/Users/gino/kaggle/fast-furious/gitHub/fast-furious/competitions/facebook-recruiting-iv-human-or-bot/"
  } else if (type == "process") {
    base.path1 = "C:/docs/ff/gitHub/fast-furious/data_process"
    base.path2 = "/Users/gino/kaggle/fast-furious/gitHub/fast-furious/data_process/"
  } else {
    stop("unrecognized type.")
  }
  
  if (file.exists(base.path1))  {
    ret = paste0(base.path1,"/")
  } else {
    ret = base.path2
  }
  
  ret
}
sampleSubmission = as.data.frame( fread(paste(getBasePath("data") , 
                                              "sampleSubmission.csv" , sep=''))) ## outcome = 0 human 

train = as.data.frame( fread(paste(getBasePath("data") , 
                                   "train.csv" , sep=''))) 

test = as.data.frame( fread(paste(getBasePath("data") , 
                                  "test.csv" , sep='')))

bids = as.data.frame( fread(paste(getBasePath("data") , 
                                  "bids.csv" , sep='')))

train.full = merge(x = bids , y = train , by="bidder_id"  )
test.full = merge(x = bids , y = test , by="bidder_id"  )

train.full$merchandise = factor(train.full$merchandise)
train.full$device = factor(train.full$device)
train.full$ip = factor(train.full$ip)
```

# Bids & bidders

Trainset bidders/bids are strictly disjoint with testset bidders/bids.
Bidders are unique inside train/test set. 

```{r}

sum(train$bidder_id %in% test$bidder_id)
sum(test$bidder_id %in% train$bidder_id)

sum(train.full$bid_id %in% test.full$bid_id)
sum(test.full$bid_id %in% train.full$bid_id)

length(unique(train$bidder_id)) == nrow(train)
length(unique(test$bidder_id)) == nrow(test)

```

# Address / Payment account 

Trainset addresses/payment accounts are stricly disjoint with testset addresses/payment accounts.
Address/payment account are unique inside train/test sets. 

```{r}

sum(train$address %in% test$address)
sum(test$address %in% train$address)

sum(train$payment_account %in% test$payment_account)
sum(test$payment_account %in% train$payment_account)

length(unique(train$payment_account)) == nrow(train)
length(unique(test$payment_account)) == nrow(test)

length(unique(train$address)) == nrow(train)
length(unique(test$address)) == nrow(test)

```

# Ips

There are 29 bidders in the train set (and 70 in the test set) that didn't partecipate to any auction. I suppose they are human ;-)
The other 1984 bidders in the train set use, on average, 675 different ip addresses (sd 4790).
Robot bidders use on average 2387 different ip addresses (sd 11269) vs. 581 different ip addresses of human (sd 4140). 
On the other hand, the same ip address has been used on average by 1.3 different bidders (sd 2.4). 

```{r}
sum(!train$bidder_id %in% bids$bidder_id)
sum(!test$bidder_id %in% bids$bidder_id)

ee = ddply(train.full,.(bidder_id),function(x)c(ip=length(unique(x$ip))))
mean(ee$ip)
sd(ee$ip)
ddf = merge(x = ee , y = train , by="bidder_id")
ddply(ddf , .(outcome) , function(x) c(ip.mean = mean(x$ip) , ip.sd = sd(x$ip) )  )

aa = ddply(train.full,.(ip),function(x)c(bidders=length(x$bidder_id)))
bb = ddply(train.full,.(ip),function(x)c(bidders=length(unique(x$bidder_id))))
cc = ddply(train.full,.(ip),function(x)c(bidders=length(unique(x$url))))

mean(bb$bidders)
sd(bb$bidders)

uu = ddply(train.full,.(bidder_id),function(x)c(url.len=length(unique(x$url))))
mean(uu$url.len)

```

# Urls 

The 1984 bidders partecipating in auctions on average have 346 different urls (sd 2676).
Humans have on average 335 different urls while robots 544.   
On the other hand, an url is used on average by 1.03 different bidders 

```{r}
uu = ddply(train.full,.(bidder_id),function(x)c(url.len=length(unique(x$url))))
mean(uu$url.len)
sd(uu$url.len)

uu = ddply(train.full,.(bidder_id,outcome),function(x)c(url.len=length(unique(x$url))))
mean(uu[uu$outcome==0,]$url.len)
sd(uu[uu$outcome==0,]$url.len)
mean(uu[uu$outcome==1,]$url.len)
sd(uu[uu$outcome==1,]$url.len)

cc = ddply(train.full,.(url),function(x)c(bidders=length(unique(x$bidder_id))))
mean(cc$bidders)
sd(cc$bidders)
```

