---
title: "Exploratory"
author: "Gino Tesei"
date: "May 24, 2015"
output: html_document
---

```{r, echo=FALSE , results="hide" , include=FALSE}
library(caret)
library(Hmisc)
library(data.table)
library(kernlab)
library(subselect)
library(plyr)
library(binhf)
library(fBasics)
library(lattice)

getBasePath = function (type = "data") {
  ret = ""
  base.path1 = ""
  base.path2 = ""
  
  if(type == "data") {
    base.path1 = "C:/docs/ff/gitHub/fast-furious/dataset/facebook-recruiting-iv-human-or-bot"
    base.path2 = "/Users/gino/kaggle/fast-furious/gitHub/fast-furious/dataset/facebook-recruiting-iv-human-or-bot/"
  } else if (type == "code") {
    base.path1 = "C:/docs/ff/gitHub/fast-furious/competitions/facebook-recruiting-iv-human-or-bot"
    base.path2 = "/Users/gino/kaggle/fast-furious/gitHub/fast-furious/competitions/facebook-recruiting-iv-human-or-bot/"
  } else if (type == "process") {
    base.path1 = "C:/docs/ff/gitHub/fast-furious/data_process"
    base.path2 = "/Users/gino/kaggle/fast-furious/gitHub/fast-furious/data_process/"
  } else {
    stop("unrecognized type.")
  }
  
  if (file.exists(base.path1))  {
    ret = paste0(base.path1,"/")
  } else {
    ret = base.path2
  }
  
  ret
}
sampleSubmission = as.data.frame( fread(paste(getBasePath("data") , 
                                              "sampleSubmission.csv" , sep=''))) ## outcome = 0 human 

train = as.data.frame( fread(paste(getBasePath("data") , 
                                   "train.csv" , sep=''))) 

test = as.data.frame( fread(paste(getBasePath("data") , 
                                  "test.csv" , sep='')))

bids = as.data.frame( fread(paste(getBasePath("data") , 
                                  "bids.csv" , sep='')))

train.full = merge(x = bids , y = train , by="bidder_id"  )
test.full = merge(x = bids , y = test , by="bidder_id"  )

train.full$merchandise = factor(train.full$merchandise)
train.full$device = factor(train.full$device)
```

# Outcome distribution 

Label 1 indicates a robot where value 0 indicates human. 

```{r}
table(train$outcome)/nrow(train)
```

## Outcome distribution by merchandise

* **Never robots** in **auto**, **clothing**, **furniture**

* **Almost never robots** in **jewelry (0.06677237)**, **office equipment (0.0495858)**, **home goods (0.04806178)**   

* **Middle cases** in  **sporting goods (0.2451847)**, **mobile (0.2135432)**, **books and music (0.1098813)**

* **Most robots** in **computers (1.198705)**

where in brackets is reported robots/human ratio. 


```{r}
histogram(~ train.full$outcome | train.full$merchandise)
ddply(train.full , .(outcome)  , function(x)  c(  merchandise = table(x$merchandise) )   )
```


## Outcome distribution by country

```{r}
histogram(~ train.full$outcome | train.full$country)
d = ddply(train.full , .(outcome)  , function(x)  c(  country = table(x$country) )   )
d[3,] = (d[2,]/d[1,])
d[3,1] = 'ratio'
head(sort(d[3,2:ncol(d)] , decreasing = T) , 30)
tail(sort(d[3,2:ncol(d)] , decreasing = T) , 30)
```

## Outcome distribution by device 

```{r}
histogram(~ train.full$outcome | train.full$device)
dd = ddply(train.full , .(outcome)  , function(x)  c(  device = table(x$device) )   )
dds = sort(dd[1,] , decreasing = T) 
dds[,1:30]
dds[,(ncol(dds)-30):ncol(dds)]
```

## Miscellaneous

```{r}
sum(train$bidder_id %in% test$bidder_id)
sum(train$payment_account %in% test$payment_account)
sum(train$address %in% test$address)
sum(test.full$auction %in% train.full$auction)
sum(! test.full$auction %in% train.full$auction)
sum(train.full$bid_id %in% test.full$bid_id)
sum(train.full$ip %in% test.full$ip)
sum(train.full$url %in% test.full$url)
```